---
title: "Factor analysis and other clustering methods"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

```{r echo = FALSE, results='hide', message=FALSE, comment=FALSE, warning=FALSE}
## load libraries
require(foreign)
require(tidyverse)
require(ReIns)
require("poweRlaw")
require(lavaan)
```

```{r, echo=FALSE}
dfPCL5 <- read.spss('/home/or/Documents/pcl5_vaData/pcl5_100k/PCL5_130K_FY19.sav', to.data.frame = TRUE)

dfPCL5 <- dfPCL5[,5:24]
dfPCL5 <- na.omit(dfPCL5)
```

```{r echo=FALSE}
dfPCL5_toBinarize <- dfPCL5
for (i in 1:20){
  # becuase under 10 its PCLN01 (02,03) need to add zero so run if
  if (i<10){
  nam <- paste("PCLN0", i, sep = "")  
  }
  else {
    nam <- paste("PCLN", i, sep = "")
  }
  pcl <- paste("PCL", i, sep = "")
  dfPCL5_toBinarize[pcl] <- dplyr::case_when(dfPCL5_toBinarize[nam]<=1 ~ 'NO', dfPCL5_toBinarize[nam]>=2 ~ 'YES')
  
}

# create specific dataframe
dfbigPCL5_Binary2 <- data_frame(dfPCL5_toBinarize[21:40])
countFreqBigPCL5 <- plyr::count(dfbigPCL5_Binary2)
```

## Factor analysis
- Here we try to present the effect of the power law distribution on factor analysis
- We also show it in other graphical manners 

```{r}
# Create new data frame
# join the two so we have frequency next to each individual
dataxCon <- dplyr::full_join(dfPCL5_toBinarize, countFreqBigPCL5, by = c("PCL1" = "PCL1", "PCL2" = "PCL2", "PCL3" = "PCL3", "PCL4" = "PCL4", "PCL5" = "PCL5", "PCL6" = "PCL6", "PCL7" = "PCL7", "PCL8" = "PCL8", "PCL9" = "PCL9", "PCL10" = "PCL10", "PCL11" = "PCL11", "PCL12" = "PCL12", 
 "PCL13" = "PCL13", "PCL14" = "PCL14", "PCL15" = "PCL15", "PCL16" = "PCL16", "PCL17" = "PCL17", "PCL18" = "PCL18", "PCL19" = "PCL19", "PCL20" = "PCL20"))

```

```{r}
dsm5Fac_Con <- '
      Int =~ PCLN01 + PCLN02 + PCLN03 + PCLN04 + PCLN05
      Av =~ PCLN06 + PCLN07
      An =~ PCLN08 + PCLN09 + PCLN10 + PCLN11 + PCLN12 + PCLN13 + PCLN14 
      Hyper =~ PCLN15 + PCLN16 + PCLN17 + PCLN18 + PCLN19 + PCLN20'

```

```{r}
model_Factor <- dsm5Fac_Con

iterateCFA <- function(data, n_itr, sample_size, freqCUT) {
  results_top = matrix(nrow = n_itr, ncol = 7)
  results_low = matrix(nrow = n_itr, ncol = 7)
  RES <- list()
  for (i in 1:n_itr) {
    dfSamp <- slice_sample(data, n = sample_size) 
    dataSAMP_top <- filter(dfSamp, freq > freqCUT)
    dataSAMP_low  <- filter(dfSamp, freq <= freqCUT)  
    factor_top <- cfa(model_Factor, data = dataSAMP_top[,4:19], estimator = "WLSMV")
    factor_low <- cfa(model_Factor, data = dataSAMP_low[,4:19], estimator = "WLSMV")
    results_top[i,1:6] <- fitMeasures(factor_top, c("chisq","df","pvalue","srmr","cfi","rmsea"))
    results_low[i,1:6] <- fitMeasures(factor_low, c("chisq","df","pvalue","srmr","cfi","rmsea"))
    results_top[i,7] <- nrow(dataSAMP_top) 
    results_low[i,7] <- nrow(dataSAMP_low) 
    }
  RES$results_top_df <- data.frame(chisq = results_top[,1], df = results_top[,2], pvalue = results_top[,3], 
                            srmr = results_top[,4], cfi = results_top[,5], rmsea = results_top[,6], subsamplesize = results_top[,7])
  RES$results_low_df <- data.frame(chisq = results_low[,1], df = results_low[,2], pvalue = results_low[,3], 
                              srmr = results_low[,4], cfi = results_low[,5], rmsea = results_low[,6], subsamplesize = results_low[,7])
  return(RES)
    
  }
  
  
```
### Sample size effects

```{r}

size_top <- c()
size_low <- c()
size_top_sd <- c()
size_low_sd <- c()

CFI_TOP <- c()
CFI_sd_TOP <- c()
CFI_LOW <- c()
CFI_sd_LOW <- c()

rmsea_TOP <- c()
rmsea_sd_TOP <- c()
rmsea_LOW <- c()
rmsea_sd_LOW <- c()

p_TOP <- c()
p_sd_TOP <- c()
p_LOW <- c()
p_sd_LOW <- c()

srmr_TOP <- c()
srmr_sd_TOP <- c()
srmr_LOW <- c()
srmr_sd_LOW <- c()

FinRES <- list()


### Set dataset for loops
datax <- dataFREQ <- datax #last equals to your dataset

# Determine frequency for splits
Splits <- c()
freq_2080 <- dataFREQ %>% 
  arrange(freq) %>% 
  slice_tail(prop = 0.2)
Splits[1] <- min(freq_2080$freq)

freq_5050 <- dataFREQ %>% 
  arrange(freq) %>% 
  slice_tail(prop = 0.5)
Splits[2] <- min(freq_5050$freq)

freq_8020 <- dataFREQ %>% 
  arrange(freq) %>% 
  slice_tail(prop = 0.8)
Splits[3] <- min(freq_8020$freq)

# set parameters
size_n <- c(250, 500, 750, 1000, 1500, 2000, 5000, 10000)
nitr <- 500

# Run loops
for (i in 1:length(size_n)){
    test <- iterateCFA(datax, n_itr = nitr, freqCUT = Splits[2], sample_size = size_n[i])
    
    size_top[i] <- median(test$results_top_df$subsamplesize)
    size_top_sd[i] <- sd(test$results_top_df$subsamplesize)
    size_low[i] <- median(test$results_low_df$subsamplesize)
    size_low_sd[i] <- sd(test$results_low_df$subsamplesize)
    
    CFI_TOP[i] <- median(test$results_top_df$cfi)
    CFI_sd_TOP[i] <- sd(test$results_top_df$cfi)
    CFI_LOW[i] <- median(test$results_low_df$cfi)
    CFI_sd_LOW [i] <- sd(test$results_low_df$cfi)
  
    rmsea_TOP[i] <- median(test$results_top_df$rmsea)
    rmsea_sd_TOP[i] <- sd(test$results_top_df$rmsea)
    rmsea_LOW[i] <- median(test$results_low_df$rmsea)
    rmsea_sd_LOW[i] <- sd(test$results_low_df$rmsea)
    
    p_TOP[i] <- median(test$results_top_df$p)
    p_sd_TOP[i] <- sd(test$results_top_df$p)
    p_LOW[i] <- median(test$results_low_df$p)
    p_sd_LOW[i] <- sd(test$results_low_df$p)
    
    srmr_TOP[i] <- median(test$results_top_df$srmr)
    srmr_sd_TOP[i] <- sd(test$results_top_df$srmr)
    srmr_LOW[i] <- median(test$results_low_df$srmr)
    srmr_sd_LOW[i] <- sd(test$results_low_df$srmr)
    }   
d_top <- data.frame(size_n, size_top, size_top_sd,
                      CFI_TOP, CFI_sd_TOP,
                      rmsea_TOP,rmsea_sd_TOP,
                      p_TOP, p_sd_TOP,
                      srmr_TOP,srmr_sd_TOP)
  
d_low <- data.frame(size_n, size_low, size_low_sd,
                    CFI_LOW, CFI_sd_LOW,
                    rmsea_LOW, rmsea_sd_LOW, 
                    p_LOW, p_sd_LOW, 
                    srmr_LOW, srmr_sd_LOW)
  
d_top$group <- c("top")
d_low$group <- c("low ")
colnames(d_low) <- colnames(d_top)
dx <- rbind(d_top, d_low)

```

```{r}
#Plot CFI
ggplot(data=dx, aes(y=CFI_TOP, x=size_n, colour=group, group=group)) +
  geom_point(size = 1.5) + 
  geom_line() +
  geom_errorbar(aes(ymin=CFI_TOP-CFI_sd_TOP, ymax=CFI_TOP+CFI_sd_TOP), width=.5, colour="black") +
  xlab("Sample size") +
  ylab("CFI") +
  scale_y_continuous(limits = c(0.80, 1.0)) +
  geom_hline(yintercept=0.95, linetype="dashed", size= 0.3) +
  theme_minimal()


# Plot RMSEA
ggplot(data=dx, aes(y=rmsea_TOP, x=size_n, colour=group, group=group)) +
  geom_point(size = 1.5) + 
  geom_line() +
  geom_errorbar(aes(ymin=rmsea_TOP-rmsea_sd_TOP, ymax=rmsea_TOP+rmsea_sd_TOP), width=.5, colour="black") +
  xlab("Sample size") +
  ylab("rmsea") +
  scale_y_continuous(limits = c(0.0, 0.15)) +
  geom_hline(yintercept=0.05, linetype="dashed", size= 0.3) +
  geom_hline(yintercept=0.08, linetype="dashed", size= 0.3) +
  theme_minimal()

# Plot SRMR
ggplot(data=dx, aes(y=srmr_TOP, x=size_n, colour=group, group=group)) +
  geom_point(size = 1.5) + 
  geom_line() +
  geom_errorbar(aes(ymin=srmr_TOP-srmr_sd_TOP, ymax=srmr_TOP+srmr_sd_TOP), width=.5, colour="black") +
  xlab("Sample size") +
  ylab("srmr") +
  scale_y_continuous(limits = c(0.0, 0.15)) +
  geom_hline(yintercept=0.1, linetype="dashed", size= 0.3) +
  theme_minimal()
  
```

### 5. EFA 

``{r}
#Installing the Psych package and loading it
install.packages("psych")
library(psych)
iterateEFA <- function(data, n_itr, sample_size, freqCUT) {
  results_top = matrix(nrow = n_itr, ncol = 2)
  results_low = matrix(nrow = n_itr, ncol = 2)
  RES_EFA <- list()
  for (i in 1:n_itr) {
    dfSamp <- slice_sample(data, n = sample_size) 
    dataSAMP_top <- filter(dfSamp, freq > freqCUT)
    dataSAMP_low  <- filter(dfSamp, freq <= freqCUT)  
    dataSAMP_top_cor <- cor(dataSAMP_top[,4:19])
    dataSAMP_low_cor <- cor(dataSAMP_low[,4:19])
    results_top[i,2] <- nrow(dataSAMP_top) 
    results_low[i,2] <- nrow(dataSAMP_low) 
    items.parallel <- fa.parallel(dataSAMP_top_cor, fa = "fa", n.obs = nrow(dataSAMP_top_cor))
    items.parallel <- fa.parallel(dataSAMP_low_cor, fa = "fa", n.obs = nrow(dataSAMP_top_cor))
    results_top[i,1] <- items.parallel$nfact
    results_low[i,1] <- items.parallel$nfact
  }
  RES_EFA$results_top_df <- data.frame("Number Factors" = results_top[,1], "sample size" = results_top[,2])
  RES_EFA$results_low_df <- data.frame("Number Factors" = results_low[,1], "sample size" = results_low[,2])
  
  return(RES_EFA)
  
}


```

```{r}
### Set dataset for loops
datax <- dataFREQ <- datax #last equals to your dataset

# Determine frequency for splits
Splits <- c()
freq_2080 <- dataFREQ %>% 
  arrange(freq) %>% 
  slice_tail(prop = 0.2)
Splits[1] <- min(freq_2080$freq)

freq_5050 <- dataFREQ %>% 
  arrange(freq) %>% 
  slice_tail(prop = 0.5)
Splits[2] <- min(freq_5050$freq)

freq_8020 <- dataFREQ %>% 
  arrange(freq) %>% 
  slice_tail(prop = 0.8)
Splits[3] <- min(freq_8020$freq)

size_n <- c(100, 250, 500, 750, 1000, 1500, 2000, 3000, 30000)
n_itr <- 300

numberfact_top <- c()
numberfact_sd_top <- c()
size_top <- c()
size_top_sd <- c()
numberfact_low <- c()
numberfact_sd_low <- c()
size_low <- c()
size_low_sd <- c()

for (i in 1:length(size_n)){
  EFA_test <- iterateEFA(datax, n_itr, freqCUT = Splits[2], sample_size = size_n[i]) 
  
  numberfact_top[i] <- median(EFA_test$results_top_df$Number.Factors)
  numberfact_sd_top[i] <- sd(EFA_test$results_top_df$Number.Factors)
  
  size_top[i] <- median(EFA_test$results_top_df$sample.size)
  size_top_sd[i] <- sd(EFA_test$results_top_df$sample.size)
  
  numberfact_low[i] <- median(EFA_test$results_low_df$Number.Factors)
  numberfact_sd_low[i] <- sd(EFA_test$results_low_df$Number.Factors)
  
  size_low[i] <- median(EFA_test$results_low_df$sample.size)
  size_low_sd[i] <- sd(EFA_test$results_low_df$sample.size)
}

d_EFA_above <- data.frame(size_n, size_top, size_top_sd,
                    numberfact_top, numberfact_sd_top)

d_EFA_below <- data.frame(size_n, size_low, size_low_sd,
                    numberfact_low, numberfact_sd_low)

d_EFA_above$group <- c("top")
d_EFA_below$group <- c("low")
colnames(d_EFA_below) <- colnames(d_EFA_above)
dx_EFA <- rbind(d_EFA_above, d_EFA_below)
```

```{r}

# Plot factors
ggplot(data=dx_EFA, aes(y=numberfact_top, x=size_n, colour=group, group=group)) +
  geom_errorbar(aes(ymin=numberfact_top-numberfact_sd_top, ymax=numberfact_top+numberfact_sd_top), colour="black") +
  geom_point(size = 1.5) + 
  geom_line() +
  xlab("Sample size") +
  ylab("Median number of factors") +
  scale_y_continuous(limits = c(0.0, 6)) +
  theme_minimal()
```
