---
title: "Symptoms of PTSD follow a power law distribution"
author: "Or Duek & Tobias Spiller"
output:
  html_document:
    df_print: paged
---

# Power law (scale-free)
- This work assess fit of power-law to different data-sets in mental health
- We fit three different distributions for each set of data:
  1. Power-law
  2. log-normal
  3. exponential
  

```{r echo = FALSE, results='hide', message=FALSE, comment=FALSE, warning=FALSE}
## load libraries
require(foreign)
require(tidyverse)
require(ReIns)
require("poweRlaw")
require(lavaan)
```

### First data-set 30k PTSD patients
- For each data set we:
  1. Load and clean data
  2. Binarize (?)
  3. Calculate profiles (combinations of symptoms) and their frequencies (how many patients per profile)
  
```{r}
dfPCL5 <- read.spss('/home/or/Documents/pcl5_vaData/pcl5_100k/PCL5_130K_FY19.sav', to.data.frame = TRUE)

dfPCL5 <- dfPCL5[,5:24]
dfPCL5 <- na.omit(dfPCL5)
```

```{r echo=FALSE}
dfPCL5_toBinarize <- dfPCL5
for (i in 1:20){
  # becuase under 10 its PCLN01 (02,03) need to add zero so run if
  if (i<10){
  nam <- paste("PCLN0", i, sep = "")  
  }
  else {
    nam <- paste("PCLN", i, sep = "")
  }
  pcl <- paste("PCL", i, sep = "")
  dfPCL5_toBinarize[pcl] <- dplyr::case_when(dfPCL5_toBinarize[nam]<=1 ~ 'NO', dfPCL5_toBinarize[nam]>=2 ~ 'YES')
  
}

# create specific dataframe
dfbigPCL5_Binary2 <- data_frame(dfPCL5_toBinarize[21:40])
countFreqBigPCL5 <- plyr::count(dfbigPCL5_Binary2)
```

### A total of 8174 combinations. 
- Describe frequencies

```{r}
## count how many with only one
sum(countFreqBigPCL5$freq==1)
# percentage
sum(countFreqBigPCL5$freq==1) / nrow(countFreqBigPCL5)
# count more than one less than ten
sum(countFreqBigPCL5$freq<10) - sum(countFreqBigPCL5$freq==1)
max(countFreqBigPCL5$freq)
max(countFreqBigPCL5$freq) / nrow(dfbigPCL5_Binary2)
```

- Plot the top 50 profiles

```{r, echo=FALSE}
freq5 <- countFreqBigPCL5[order(countFreqBigPCL5$freq, decreasing = TRUE),]
freq5_top50 <- freq5[1:50,]
x <- as.factor(1:nrow(freq5_top50))
ggplot(freq5_top50, aes(x=as.factor(x),y=freq)) + geom_bar(stat="identity") + xlab("Combination Of Symptoms") + theme_minimal()
```

### Test distributions
#### Power-law

```{r, echo=FALSE}
m_pl = displ$new(countFreqBigPCL5$freq)
est = estimate_xmin(m_pl) ## Estimate parameters
m_pl$setXmin(est) ## Update object
#plot(m_pl)
```

* Use bootstrap to test significant level 

```{r, echo=FALSE}
bs_p = bootstrap_p(m_pl, no_of_sims = 1000, threads = 10)
plot(bs_p)
```

#### Log-Normal

```{r}
m_ln = dislnorm$new(countFreqBigPCL5$freq)
est_ln = estimate_xmin(m_ln) ## Estimate parameters
m_ln$setXmin(est_ln)
#plot(m_ln)
```

```{r, echo=FALSE}
bs_ln = bootstrap(m_ln, no_of_sims = 1000, threads = 10)
plot(bs_ln)
```

#### Exponential

```{r}
m_exp = disexp$new(countFreqBigPCL5$freq)
est_exp = estimate_xmin(m_exp) ## Estimate parameters
m_exp$setXmin(est_exp)
```
```{r, echo=FALSE}
bs_exp = bootstrap_p(m_exp, no_of_sims = 1000, threads = 10)
plot(bs_exp)
```

### Compare distributions
- Lognormal and power-law

```{r}
m_ln$setXmin(m_pl$getXmin())
est2 = estimate_pars(m_ln)
m_ln$setPars(est2$pars)

comp = compare_distributions(m_pl, m_ln) # Test the two distributions
comp$p_two_sided #not sig: inconclusive
comp$p_one_sided #not sig: inconclusive
```
- results are inconclusive. 
- Compare to exponential

```{r}
m_exp$setXmin(m_pl$getXmin())
est3 = estimate_pars(m_exp)
m_exp$setPars(est3$pars)

comp2 = compare_distributions(m_pl, m_exp) # Test the two distributions
comp2$p_two_sided #not sig: inconclusive
comp2$p_one_sided #not sig: inconclusive
```

- Still inconclusive, but seems closer to significant differece.

## Move on to second database - DSM-IV

```{r echo = FALSE, results='hide'}
# load data set
source('/home/or/Documents/va_data/readData.r')

# all patientes with PTSD and PCLTOT
pclAll <- dplyr::filter(vaDatclean, !is.na(BPCLTOT))
# plot pcl total score 
hist(pclAll$BPCLTOT)
# we have a minimum of 2 - so we have some NAs - let remove them
pclAll_Nas <- filter(pclAll, BPCLTOT <=16)
# total of 20 subjects with 16 or less in PCL (i.e. at least one missing variable)
# we can remove them from analysis
pclAll <- filter(pclAll, BPCLTOT >=17)
# 159577 patients
#pclNetwork <- pclNoNa # just medicated
pclNetwork <- pclAll
nrow(pclNetwork)
hist(pclNetwork$BPCLTOT)
# take just pcl
pclItems <- dplyr::select(pclAll, starts_with("PCL"))
pclItems_noCluster <- dplyr::select(pclItems, -PCLFY, -PCLSURVEYDATE, -PCLRAWSCORE)
nrow(pclItems_noCluster)
pclItems_noCluster <- na.omit(pclItems_noCluster)
nrow(pclItems_noCluster)

# Binarization
dfPCL4 <- pclItems_noCluster
# using PCL-4 recommended threshold is 3+ (not 2+ as in PCL5)
for (i in 1:17){
  nam = paste("PCL", i, sep = "")
  pcl <- paste("PCLB", i, sep = "")
  dfPCL4[pcl] <- dplyr::case_when(dfPCL4[nam]<=2 ~ 'NO', dfPCL4[nam]>=3 ~ 'YES')
  
}

# create specific dataframe
id = 1:nrow(dfPCL4)
dfPCL4_Binary <- data_frame(id, dfPCL4[18:34])
countFreq_pcl4 <- plyr::count(dfPCL4_Binary[,-1])

```


- Frequencies:

```{r}
## count how many with only one
sum(countFreq_pcl4$freq==1)
# percentage
sum(countFreq_pcl4$freq==1) / nrow(countFreq_pcl4)
# count more than one less than ten
sum(countFreq_pcl4$freq<10) - sum(countFreq_pcl4$freq==1)
max(countFreq_pcl4$freq)
max(countFreq_pcl4$freq) / nrow(dfPCL4_Binary)
```

- Plot the top 50 profiles

```{r, echo=FALSE}
freq4 <- countFreq_pcl4[order(countFreq_pcl4$freq, decreasing = TRUE),]
freq4_top50 <- freq4[1:50,]
x <- as.factor(1:nrow(freq4_top50))
ggplot(freq4_top50, aes(x=as.factor(x),y=freq)) + geom_bar(stat="identity") + xlab("Combination Of Symptoms") + theme_minimal()
```

```{r}
m4_pl = displ$new(countFreq_pcl4$freq)
est_m4 = estimate_xmin(m4_pl) ## Estimate parameters
m4_pl$setXmin(est_m4) ## Update object
```

* Use bootstrap to test significant level 

```{r, echo=FALSE}
bs4_p = bootstrap_p(m4_pl, no_of_sims = 1000, threads = 10)
plot(bs4_p)
```

#### Log-Normal

```{r}
m4_ln = dislnorm$new(countFreq_pcl4$freq)
est4_ln = estimate_xmin(m4_ln) ## Estimate parameters
m4_ln$setXmin(est4_ln)
#plot(m_ln)
```

```{r, echo=FALSE}
bs4_ln = bootstrap(m4_ln, no_of_sims = 1000, threads = 10)
plot(bs4_ln)
```

#### Exponential

```{r}
m4_exp = disexp$new(countFreq_pcl4$freq)
est4_exp = estimate_xmin(m4_exp) ## Estimate parameters
m4_exp$setXmin(est4_exp)
```
```{r, echo=FALSE}
bs_exp = bootstrap_p(m_exp, no_of_sims = 1000, threads = 10)
plot(bs_exp)
```

### Compare distributions
- Lognormal and power-law

```{r}
m4_ln$setXmin(m4_pl$getXmin())
est4_comp = estimate_pars(m4_ln)
m4_ln$setPars(est4_comp$pars)

comp = compare_distributions(m4_pl, m4_ln) # Test the two distributions
comp$p_two_sided #not sig: inconclusive
comp$p_one_sided #not sig: inconclusive
```
- results are inconclusive. 
- Compare to exponential

```{r}
m4_exp$setXmin(m4_pl$getXmin())
est4_exp = estimate_pars(m4_exp)
m4_exp$setPars(est4_exp$pars)

comp3 = compare_distributions(m4_pl, m4_exp) # Test the two distributions
comp3$p_two_sided #not sig: inconclusive
comp3$p_one_sided #not sig: inconclusive
```
